---
title: "Challenge_7_W"
author: "Kevin Flores Gonzalez"
format:
  html: 
    code-fold: true
    code-summary: "Show the code"
    theme: lux
    toc: true
    self-contained: true
    code-tools: true
editor: visual
execute:
  echo: true
  error: true
  messages: false
---

# Challenge 7: Incorporating Multiple Inputs 

A frequently used measurement for fish health is a \"condition index.\" [(Wikepedia article)](https://en.wikipedia.org/wiki/Condition_index_in_fish) The following simple equation can be used to calculate the approximate condition index of a fish:

$$
condition \ index = \frac{weight}{length^3} \times  100
$$

### Loading in the Dataset and Libraries 

```{r}
library(tidyverse)
library(dplyr)
bf_fish <- read_csv(here::here
                    ("Week 7", "Lab 7", "Blackfootfish.csv"))
```

## 4.1: Part 1

There are specific units required for the calculation of a condition index. Length must be in millimeters, and weight must be in grams. Inspect the `length` and `weight` variables to decide if you believe these are the correct units associated with these measurements---this will likely require Googling what \"typical\" measurements of trout are.

### **Replacing Impossible Measurements with `NA`s**

Based on your research, write function(s) to handle the unlikely / impossible measurements included in the dataset. Your function(s) should accept three inputs (1) a vector of measurements, (2) the minimum value you believe is \"reasonable,\" and (3) the maximum value you believe is \"reasonable.\" If a value falls outside these bounds, you should replace it with an `NA`.

Use your function to modify the `length` and `weight` columns of the `BlackfootFish` dataset, removing values you believe are \"unreasonable.\"

```{r}
Replacing_Impossible_Measurements <- function(vec,
                                              min_val,
                                              max_val){
  stopifnot(is.numeric(vec))
  case_when(
    {{vec}} > max_val ~ NA_real_,
    {{vec}} < min_val ~ NA_real_,
    TRUE ~ {{vec}}
  )
}
```

```{r}
bf_fish <- bf_fish |> 
  mutate(
        length = Replacing_Impossible_Measurements(length, 
                                                   120, 
                                                   300),
        weight =  Replacing_Impossible_Measurements(weight, 
                                                    70, 
                                                    200)
  )
```

## 4.2: Part 2

Write a function which calculates the condition index of a fish, given inputs of weight and length.

```{r}
Condition_Index = function(data, 
                           len_val, 
                           wt_val){
  data <- data |> 
    na.omit() |> 
    mutate(`Condition Index` = ({{wt_val}} / ({{len_val}}^3)) * 100)
  return(data)
}
```

## 4.3: Part 3

Make a thoughtful visualization of how fish conditions have varied over the duration of this study.

```{r}
bf_fish_2 <- bf_fish |> 
  Condition_Index(len_val = length, 
                  wt_val = weight) |> 
  mutate(Named_Trips = ifelse(trip == 1, 
                         "Trip 1", 
                         "Trip 2")) |> 
  group_by(Named_Trips, 
           year, 
           species, 
           section) |> 
  summarize(`Condition Index` = mean(`Condition Index`)) 
```

```{r}
ggplot(data = bf_fish_2, mapping = aes(x = year, 
                                       y = `Condition Index`)) +
  geom_line() +
  facet_grid(cols = vars(species), 
             rows = vars(Named_Trips)) + 
  labs(x = "Year",
       title = "Mean Condition Index For Each Year By Trip And Species",
       y = "")
```
