---
title: "Challenge_3_W"
author: "Kevin Flores Gonzalez"
format:
  html: 
    code-fold: true
    code-summary: "Show the code"
    theme: lux
    toc: true
    self-contained: true
    code-tools: true
editor: visual
execute:
  echo: true
  error: true
  messages: false
---

## Loading in Data & Declaring Libraries

```{r}

#| message: FALSE
library("tidyverse")
hiphop <- read_csv(here::here
                    ("Week 3", "Lab 3", "hiphop.csv"))
```

## Male versus Female

```{r}
gender_hiphop <- hiphop |> 
  group_by(sex) |> 
  distinct(subj, .keep_all = TRUE) |> 
  summarize(across(intl:unclassifiable, mean)) |> 
  summarize(across(intl:unclassifiable, diff))
gender_hiphop
```

Values that are negative indicate that Males like it more, while values that are positive indicate that Females like it more.

The genre that differentiates Male and Female groups the most is rock. Rock has the highest value in the tibble above.

## White versus Non-White

```{r}
ethnicity_hiphop <- hiphop |> 
  mutate(ethnic_categorical = ifelse(ethnic == "white", 
                         "white", 
                         "non-white")) |> 
  group_by(ethnic_categorical) |> 
  distinct(subj, .keep_all = TRUE) |>
  summarize(across(intl:unclassifiable, mean)) |> 
  summarize(across(intl:unclassifiable, diff))
ethnicity_hiphop
```

Values that are negative indicate Non-White preference while values that positive indicate White preference.

Intl: Higher for Non-Whtie

Rock:Higher for White

Alternative: Higher for White

Hip Hop: Higher for White

## Concerns over Data Collection Method

-   All participants of this study were collected form a single university. This limits the usefulness of the data to only that university since it is not Representative of other universities.

-   All students were students form Minnesota, leading to more issues as students in Minnesota likely differ from students in other states.

-   The study excludes Black individuals while using black stereotypes to gauge different peoples understanding of Black culture. The study as a whole is unethical and perpetuates stereotypes.
