---
title: "Lab_4_W"
author: "Kevin Flores Gonzalez"
format:
  html: 
    code-fold: true
    code-summary: "Show the code"
    theme: lux
    toc: true
    self-contained: true
    code-tools: true
editor: visual
execute:
  echo: true
  error: true
  messages: false
---

## Lab 4: Avocados

For this analysis we are going to be looking at avocado prices.

The avocado data contains various statistics on the sales, volume, and production of avocados in various regions.

## Loading in the Data & Libraries

```{r}
#| message: false
library(tidyverse)
library(dplyr)
avocado <- read_csv(here::here
                   ("Week 4", "Lab 4", "avocado.csv"))
```

## Cleaning the Data

In the following code, the avocado dataset was cleaned

```{r}
library(dplyr)
avocado_major_regions <- data.frame(
  region = c("Midsouth", 
             "Northeast", 
             "Plains", 
             "Southeast", 
             "West", 
             "SouthCentral", 
             "GreatLakes", 
             "WestTexNewMexico" )
)

avocado_cali_regions <- data.frame(
  region = c("SanFrancisco", 
             "LosAngeles", 
             "SanDiego",
             "Sacramento")
  )

avocado_states <- data.frame(
  region = c("California", 
             "NewYork", 
             "SouthCarolina")
  )

avocado <- avocado |> 
  rename("small" = `4046`, 
         "large" = `4225`, 
         "extralarge" = `4770`) |> 
  separate(col=Date, into=c('Year', 
                            'Month', 
                            "Day"), sep='-') 
```

## Region that sold the most Hass avocados in 2017

Based on the analysis below we can conclude that in 2017, the region that sold the most Hass avocados was the "West" region.

```{r}
avocado |>
  semi_join(y = avocado_major_regions, 
            by = "region") |> 
  filter(type == "organic", 
         Year == 2017) |> 
  group_by(region) |> 
  summarise(sum_small_total_volume = sum(small))
```

## Highest Volume of Avocado Sales per Month

Based on the analysis below we can conclude that the region with the highest volume of avocado sales per month is February.

```{r}
avocado |> 
  select(Month, `Total Volume`) |> 
  group_by(Month) |> 
  summarise(`Total Per Month` = sum(`Total Volume`)) |> 
  slice_max(`Total Per Month`)
```

## Metro Area Regions that sold the Most Total Avocados

Based on the analysis below we can conclude that the top five regions that sold the most avocados are:

1\) Phoenix/Tucson

2\) Houston

3\) Denver

4\) Dallas Ft Worth

5\) BailtimoreWashington

```{r}

Metro_Avocado_Only <- avocado |> 
  anti_join(avocado_major_regions, 
            by = "region") |> 
  anti_join(avocado_cali_regions, 
            by = "region") |> 
  anti_join(avocado_states, 
            by = "region") |> 
  filter(region != "TotalUS")

Metro_Top5 <- Metro_Avocado_Only |> 
  group_by(region) |> 
  summarize("Sum Total Volume" = sum(`Total Volume`)) |> 
  arrange(desc(`Sum Total Volume`)) |> 
  slice_head(n = 5)

Metro_Avocado_Only |> 
  semi_join(Metro_Top5, 
            by = "region") |> 
  ggplot(mapping = aes(x = `Total Volume`, 
                       y = region))+
  geom_boxplot()+
  labs(x = "Total Volume Of Avocado Sales", 
       y = "Top Five Metro Areas")
```

## Reshaping Data: California Regions Only

```{r}
#| message: FALSE
#| warning: FALSE
organic_vs_conventional_avocado <- avocado |>
  semi_join(avocado_cali_regions, 
            by = "region") |> 
  group_by(region, type) |> 
  summarise(avg_avocado_price = mean(AveragePrice)) |> 
  pivot_wider(names_from = "type", 
              values_from = "avg_avocado_price") |> 
  mutate(`Difference In Avg Price For Conventional and Organic` = organic - conventional) 

ggplot(organic_vs_conventional_avocado, 
       aes(x = region,
    y = `Difference In Avg Price For Conventional and Organic`))+
geom_point()


```

San Francisco is the region where the price of organic versus conventional avocados are most different.

### Plot of Hass Avocado Sales That are Small, Large, or Extra Large; Conventional vs. Organic.

```{r}
proportion_avocado_data <- avocado |>
  semi_join(avocado_cali_regions, 
            by = "region") |> 
  pivot_longer("extralarge":"small",
               names_to = "Avocado Size",
               values_to = "Number Of Avocados") |> 
  mutate(Proportion = `Number Of Avocados` / mean(`Total Volume`))
proportion_avocado_data |> 
  ggplot(mapping = aes(x = region, 
                       y = Proportion, 
                       fill = `Avocado Size`))+
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity") +
  labs(x = "Regions of CA", 
       y = "Proportion of Mean Avocados Sold") + 
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  facet_wrap(~type)
```
