---
title: "Lab_9"
author: "Kevin Flores Gonzalez"
format:
  html: 
    code-fold: true
    code-summary: "Show the code"
    theme: lux
    toc: true
    self-contained: true
    code-tools: true
editor: visual
execute:
  echo: true
  error: true
  messages: false
---

# Lab 9

## 1: The Data

Our dataset(s) in this lab concerns baby names and their popularity over time. At [this link](https://www.dropbox.com/sh/czm61yzmvzl6xfb/AABEFdpy02aNRhZ_vcxrVAQSa?dl=0), you can find the names for ALL 50 states, in separate datasets organized by first letter. For each year, and for each name with at least 50 recorded babies born, we are given the counts of how many babies had that name.

## Importing the Dataset and Libraries

```{r}
#| message: false
library(tidyverse)
library(here)
StateNamesA <- read_csv(here::here("Week 9", 
                                    "Lab & Challenge", 
                                    "StateNames_A.csv"))
```

## 2: Is my name not any cool more?

Let's take a look at how the name "Allison" has changed over time. As my name begins with "A", you should download the `StateNames_A.csv` dataset from the link above.

## 3: Summarizing & Visualizing the Number of Allison's

1.  Make a summary table of the number of babies named "Allison" for each state and the sex of the baby. Specifically, the table should have the following qualities:

-   each state should be its own row

-   and each sex should have its own column

-   if there were no babies born for that combination of state & sex there should be a 0 (not an `NA`)Running Code

```{r}
StateNamesA <- StateNamesA |> 
  rename("Sex at Birth" = "Gender") 

StateNamesA |> 
  pivot_wider(
    names_from = `Sex at Birth`,
    values_from = Count) |> 
  filter(Name == "Allison") |> 
  group_by(State) |> 
  summarize(across(F:M, 
                   sum)) |> 
  mutate(across(F:M, 
                ~replace_na(.,0)))
```

2.  You should have seen in the table above that "Allison" is a name given overwhelmingly to babies assigned "female" at birth. So, create a new dataset named `allison_f` which contains **only** the babies assigned `F`emale at birth.

```{r}
allison_f <- StateNamesA |> 
  filter(`Sex at Birth` == "F",
          Name == "Allison") |> 
  group_by(Year) |> 
  summarize(Sum_Allison = sum(Count))
```

3.  Make a visualization showing how the popularity of the name "Allison" has changed over the years. To be clear, each year should have **one** observation--the total number of Allison's born that year.

```{r}
allison_f |> 
  ggplot(mapping = aes(x = Year, 
                       y = Sum_Allison)) + 
  geom_col() +
  labs(title = "Number of Females Named Allison per Year in the US",
       y = "")
```

## 4: Modeling the Number of Allison's

4.  Fit a linear model with the year as the explanatory variable, and the number of Allisons as the response. Similar to #3, each year should have **one** observation--the total number of Allisons born that year.

```{r}
Regression_Allison <- StateNamesA |> 
  filter(Name == "Allison",
        `Sex at Birth` == "F") %>%
  lm(Count ~ Year, 
     data = .)
summary(Regression_Allison)
```

5.  Write out the estimated regression equation.

    $$
    \hat{y} = 3894.36 - 1.876
    $$

6.  Plot the *residuals* of the model, that is, the actual values minus the predicted values. Comment on the residuals - do you see any patterns?

```{r}
Regression_Allison |> 
  broom::augment() |> 
  ggplot(mapping = aes(x = .fitted, 
                       y = .resid)) +
  geom_point() +
  labs(y = "Residual Values", 
       x = "Fitted Values", 
       title = "Visualization Of Residuals Vs Fitted Values")
```

7.  What do you conclude from this model? Is my name not cool anymore?

Since the vast majority of residuals lie around zero, and appear to be random it is reasonable to assume that the model has a linear relationship. The slope tells us that on average each additional year decreases the popularity of the name Allison by -1.876. The name is declineing in popularity, and if that is what we are using to decide if it "cool", the it is getting more "uncool" by the year.

## 5: Spelling by State

In middle school I was **so** upset with my parents for not naming me "Allyson". Past my pre-teen rebellion, I'm happy with my name and am impressed when baristas spell it "Allison" instead of "Alison". But I don't have it as bad as my good friend Allan!

8.  Narrow the A name dataset (downloaded previously) down to only male-assigned babies named "Allan", "Alan", or "Allen". Make a plot comparing the popularity of these names over time.

    ```{r}
    StateNamesA |> 
      filter(`Sex at Birth` == "M",
             Name %in% c("Allan", 
                         "Alan",
                         "Allen")) |>
      ggplot(mapping = aes(x = Year, 
                           y = Count, 
                           color = Name)) +
      geom_smooth() +
      labs(title = "Occurencies Of Different Spellings Of Allen Over Time")
    ```

9.  In California, Allan's spelling of his name is the least common of the three but perhaps it's not such an unusual name for his home state of Pennsylvania. Compute the total number of babies born with each spelling of "Allan" in the year 2000, in Pennsylvania and in California. Specifically, the table should have the following qualities:

-   each spelling should be its own column

-   each state should have its own row

-   a `0` (not an `NA`) should be used to represent locations where there were no instances of these names

10. Convert your total counts to overall percents. That is, what was the percent breakdown between the three spellings in CA? What about in PA?

```{r}
StateNamesA |> 
  filter(`Sex at Birth` == "M",
         Name %in% c("Allan", 
                     "Alan", 
                     "Allen")) |> 
  pivot_wider(
    names_from = Name,
    values_from = Count) |> 
  filter(
    Year == 2000,
    State %in% c("PA", 
                 "CA")) |>
  mutate(
    across(Allan:Allen, ~ replace_na(., 0)),
    AllenPercentage = round(Allen/ rowSums(across(Allen:Allan)), 2),
    AlanPercentage = round(Alan/ rowSums(across(Allen:Allan)), 2),
    AllanPercentage = round(Allan/ rowSums(across(Allen:Allan)), 2),
  ) |> 
  subset(select = -c(1,2))
```
